---
title: "MTG sim scratchpad"
author: "Jonathan Che"
date: "2024-12-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(data.tree)
require(tidyverse)
COLORS <- c("W", "U", "B", "R", "G")
```

# Overview

Instead of doing math, let's run some numbers and make some pictures to understand how to build optimal manabases in limited.
Pet project to learn about S3 classes in R (https://adv-r.hadley.nz/s3.html#s3).

# Functions

## Card class

A card is a string `[S/L][WUBRG][T]` denoting:

* `[S/L]`: spell/land
* `[WUBRG]`: color(s)
* `[T]`: enters tapped?

```{r}
new_card <- \(x = character()) {
  stopifnot(is.character(x))
  structure(x, class="card")
}

validate_card <- \(x) {
  values <- unclass(x)
  colors <- attr(x, "colors")
  tapped <- attr(x, "tapped")
  
  if(!is.character(x)) {
    stop("Card must be represented by a string")
  }
  if (!all(colors %in% COLORS)) {
    stop("Card must have colors in WUBRG")
  }
  if (!is.logical(tapped)) {
    stop("Card must either enter tapped or untapped")
  }
  
  x
}

card <- \(x = character()) {

  attr(x, "type") <- x |> 
    substr(1,1) |> 
    toupper() |> 
    switch(S="spell", F="fetch", "land")
  attr(x, "colors") <- COLORS[sapply(COLORS, \(y) grepl(y, x))]
  attr(x, "tapped") <- substr(x, nchar(x), nchar(x)) == "T"
  
  x |> 
    new_card() |> 
    validate_card()
}

print.card <- \(x) {
  cat("Class:", attr(x, "type"), "\n")
  if (attr(x, "type") == "land") cat("Colors:", attr(x, "colors"), "\n")
  if (attr(x, "tapped")) cat("Enters tapped")
}

is.card <- \(x) {
  class(x) == "card"
}
```

## Deck class

A deck is a list of cards.
By default, decks are constructed by passing in a list of cards.
Helper function `build_deck()` offers a shortcut for constructing decks, by passing in card names followed by card quantities, e.g., `d <- build_deck(W=8, U=7, WUT=2)`.

```{r}
new_deck <- \(x = list(), size = integer(), lands = integer()) {
  stopifnot(is.list(x))
  stopifnot(is.integer(size))
  stopifnot(is.integer(lands))
  
  structure(x, class="deck", size=size, lands=lands)
}

validate_deck <- \(x) {
  if (!all(sapply(x, is.card))) {
    stop("Deck must only contain cards")
  }
  
  if(attr(x, "size") < attr(x, "lands")) {
    stop("Number of lands cannot exceed deck size")
  }
  
  x
}

deck <- \(x = list(), size = length(x), lands = 0) {
  stopifnot(all(sapply(x, is.card)))
  
  # record number of lands, if x is nonempty and lands is not specified
  if (lands == 0 & size > 0) {
    lands <- x |> 
      sapply(function(x) attr(x, "type") == "land") |> 
      sum()
  }
  
  x |> 
    new_deck(size=size, lands=lands) |> 
    validate_deck()
}

print.deck <- \(x) {
  cat("Size:", attr(x, "size"), "\n")
  cat("Lands:", attr(x, "lands"), "\n")
  cat("Spells:", attr(x, "size") - attr(x, "lands"), "\n")
  cat("Cards:", paste(x, collapse=", "))
}

is.deck <- \(x) {
  class(x) == "deck"
}

summary.deck <- \(x) {
  x |> 
    as.character() |> 
    table()
}

# helper function to shortcut building a deck
#  - instead of individually  specifying each card, 
#    specify list of [card] = [count]
build_deck <- \(num_cards = 40, ...) {
  args <- list(...)
  
  # populate deck with named lands & remaining slots with spells
  vec <- rep(names(args), args)
  if (length(vec) < num_cards) {
    vec <- c(vec, rep("S", num_cards-length(vec)) )
  }

  vec |> 
    lapply(card) |> 
    deck()
}
```

## Deck functions

Functions that we can apply to decks.
Treat hands as smaller decks.

```{r}
# shuffle deck
shuffle <- \(d) {
  stopifnot(is.deck(d))
  d |> 
    sample(length(d), replace=F) |> 
    deck()
}

# draw top `size` cards from deck
draw <- \(d, size=7) {
  stopifnot(is.deck(d))
  deck(d[1:size])
}
```

## Counting functions

Functions to count colored sources.
This is where the juice of our object-oriented setup gets squeezed!

```{r}
size <- \(d) {
  stopifnot(is.deck(d))
  attr(d, "size")
}
count_spells <- \(d) {
  stopifnot(is.deck(d))
  attr(d, "size") - attr(d, "lands")
}
count_lands <- \(d) {
  stopifnot(is.deck(d))
  attr(d, "lands")
}

# count color sources in hand
count_colors <- \(d) {
  stopifnot(is.deck(d))
  
  COLORS |> 
    sapply(\(color) {
      d |> 
        sapply(\(card) {color %in% attr(card, "colors")}) |> 
        sum()
    })
}

check_hand <- \(d) {
  stopifnot(is.deck(d))
  sources <- count_colors(d)
  cat(paste(d), "\n")
  cat(count_spells(d), "spells, ", count_lands(d), "lands\n")
  cat("Color sources:\n")
  cat("W: ", sources["W"], "\n")
  cat("U: ", sources["U"], "\n")
  cat("B: ", sources["B"], "\n")
  cat("R: ", sources["R"], "\n")
  cat("G: ", sources["G"], "\n")
}
```

## Opening hand functions

Function to count number of distinct opening hands.
Requires dynamic programming.

```{r}
count_opening_hands <- \(d, n=7) {
  tab <- summary(d)
  m <- length(tab)
  # Create dp matrix: rows for categories, columns for possible sums
  dp <- matrix(0, nrow = m + 1, ncol = n + 1)
  dp[1, 1] <- 1  # Base case: 0 sum with 0 categories has 1 way
  
  # For each category
  for (i in 1:m) {
    # For each possible sum
    for (s in 0:n) {
      # For each possible value we could take from current category
      for (k in 0:min(tab[i], s)) {
        # Add number of ways to make remaining sum with previous categories
        dp[i + 1, s + 1] <- dp[i + 1, s + 1] + dp[i, s - k + 1]
      }
    }
  }
  print(dp)
  dp[m + 1, n + 1]
}
```

Function to calculate probability of given opening hand.
Multivariate hypergeometric.

```{r}

```

Function to enumerate all possible opening hands.

Maybe just do this when counting them...?

```{r}
get_opening_hands <- \(d) {
  
}
```




## Example

```{r}
d <- build_deck(W=8, U=8, WUT=1)
d

d |> shuffle() |> draw() |> check_hand()
```


# Building trees

To count number of untapped color sources by turn x...
It's a recursive process! 
If I have one untapped land and one tapped land in my opener (and don't draw), I can have:

* One untapped land T1, one untapped land T2
* No untapped lands T1, two untapped lands T2

These choices branch out!
The optimal plays to make depend on the mana value you're trying to hit, as well as your remaining deck.

## Brainstorming

Option 1: given that there are only 40 cards in a limited deck, seems pretty computationally feasible to branch out all land-playing options and store in a tree.
This might get annoying when running a bajillion simulations, but hard to tell?
Note that we don't need to care which island gets played, e.g.
Also, could we not just compute all the probabilities in closed form here...?

Option 2: devise some sort of greedy algorithm.
Issue is that this might not correspond to actual play.

Solution: do option 1, but we just get exact probabilities -- no need to simulate these!
It is annoying that 40 choose 7 is still ~18.6mil (though we get to eliminate some combinations, still on the order of 1mil) -- we'll likely still have to simulate starting hands.


## `data.tree`

```{r}
d <- build_deck(W=8, U=8, WUT=1) |> shuffle()

t <- Node$new(name = "deck", hand = d |> shuffle() |> draw())

for (card in unique(t$Get("deck"))) {
  
}
```


## Can we just compute everything in closed form?

Opening hands: in a monocolored 17-land deck, there are only 8 possible opening hands, each with probability $choose(n_l, k_l) * choose(n_s, k_s) / choose(40, 7)$.
This would be incredibly easy to enumerate.

With three card types, we end up with 36 options (assuming all card types have at least 7 cards).
As we increase the number of card types, expect a factorial explosion.

Q: is there a simple closed-form formula for enumerating the number of possible hands?
I.e., given a jar with marbles of a bunch of different colors, how many unique color combos can I draw?

A: Multivariate hypergeometric gives probability of each color combo.
Size of support of multivariate hypergeometric gets us number of options.
Expression for support looks like an optimization problem lol
https://en.wikipedia.org/wiki/Hypergeometric_distribution#Multivariate_hypergeometric_distribution

Eventually, this is going to factorially explode.
E.g., with 40 unique cards, there are $40*39*38*37*36*35*34*33 = 3e12$ starting hand options.
We might be in the computable range with most realistic scenarios, though, in which case we could just go ahead and fully enumerate everything in a single tree!


# TODO, old

* Implement "count untapped lands by turn x"
* Implement terramorphic expanse
* Implement MDFC


# TODO, new

* Probably just do things in python? Annoying to have to work with a tree package.

# Final design

A card object has the following attributes and functions [future features in brackets]

* type: land/spell/[other, e.g., creature/spell]
* color: WUBRG[V]
* tapped: T/F
* [mv: 0-16]
* [a card is hashable]

A deck object has the following attributes and functions [future features in brackets]

* deck: dictionary of cards:counts
* draw(): randomly return one card and remove it from the deck

A game object has the following attributes [future features in brackets]

* turn: turn count
* hand: list of cards
* deck: deck
* [probability: conditional probability of ending up in this state, given previous state]

A game object has the following functions [future features in brackets]

* turn: draw a card, untap tapped cards
* count_lands: count lands in hand
* count_colors: count color sources in hand

A gametree object is a tree, where each node is a game.
The first layer of nodes contains all possible opening hands.
Subsequent layers contain all possible sequences of draws.


Note: this isn't the most efficient data representation, since game states can merge.
To leave room for future work (e.g., casting spells, tempo, etc.), we'll stick with this representation.








