---
title: "MTG sim scratchpad"
author: "Jonathan Che"
date: "2024-12-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(tidyverse)
COLORS <- c("W", "U", "B", "R", "G")
```

# Overview

Instead of doing math, let's run some numbers and make some pictures to understand how to build optimal manabases in limited.
Pet project to learn about S3 classes in R (https://adv-r.hadley.nz/s3.html#s3).

# Functions

## Card class

A card is a string `[S/L][WUBRG][T]` denoting:

* `[S/L]`: spell/land
* `[WUBRG]`: color(s)
* `[T]`: enters tapped?

```{r}
new_card <- \(x = character()) {
  stopifnot(is.character(x))
  structure(x, class="card")
}

validate_card <- \(x) {
  values <- unclass(x)
  colors <- attr(x, "colors")
  tapped <- attr(x, "tapped")
  
  if(!is.character(x)) {
    stop("Card must be represented by a string")
  }
  if (!all(colors %in% COLORS)) {
    stop("Card must have colors in WUBRG")
  }
  if (!is.logical(tapped)) {
    stop("Card must either enter tapped or untapped")
  }
  
  x
}

card <- \(x = character()) {

  attr(x, "type") <- x |> 
    substr(1,1) |> 
    toupper() |> 
    switch(S="spell", F="fetch", "land")
  attr(x, "colors") <- COLORS[sapply(COLORS, \(y) grepl(y, x))]
  attr(x, "tapped") <- substr(x, nchar(x), nchar(x)) == "T"
  
  x |> 
    new_card() |> 
    validate_card()
}

print.card <- \(x) {
  cat("Class:", attr(x, "type"), "\n")
  if (attr(x, "type") == "land") cat("Colors:", attr(x, "colors"), "\n")
  if (attr(x, "tapped")) cat("Enters tapped")
}

is.card <- \(x) {
  class(x) == "card"
}
```

## Deck class

A deck is a list of cards.
By default, decks are constructed by passing in a list of cards.
Helper function `build_deck()` offers a shortcut for constructing decks, by passing in card names followed by card quantities, e.g., `d <- build_deck(W=8, U=7, WUT=2)`.

```{r}
new_deck <- \(x = list(), size = integer(), lands = integer()) {
  stopifnot(is.list(x))
  stopifnot(is.integer(size))
  stopifnot(is.integer(lands))
  
  structure(x, class="deck", size=size, lands=lands)
}

validate_deck <- \(x) {
  if (!all(sapply(x, is.card))) {
    stop("Deck must only contain cards")
  }
  
  if(attr(x, "size") < attr(x, "lands")) {
    stop("Number of lands cannot exceed deck size")
  }
  
  x
}

deck <- \(x = list(), size = length(x), lands = 0) {
  stopifnot(all(sapply(x, is.card)))
  
  # record number of lands, if x is nonempty and lands is not specified
  if (lands == 0 & size > 0) {
    lands <- x |> 
      sapply(function(x) attr(x, "type") == "land") |> 
      sum()
  }
  
  x |> 
    new_deck(size=size, lands=lands) |> 
    validate_deck()
}

print.deck <- \(x) {
  cat("Size:", attr(x, "size"), "\n")
  cat("Lands:", attr(x, "lands"), "\n")
  cat("Spells:", attr(x, "size") - attr(x, "lands"), "\n")
  cat("Cards:", paste(x, collapse=", "))
}

is.deck <- \(x) {
  class(x) == "deck"
}

# helper function to shortcut building a deck
#  - instead of individually  specifying each card, 
#    specify list of [card] = [count]
build_deck <- \(num_cards = 40, ...) {
  args <- list(...)
  
  # populate deck with named lands
  vec <- rep(lapply(names(args), card), args)

  # fill rest of deck with spells
  if (length(vec) < num_cards) {   
    vec <- c(vec, lapply(rep("S", num_cards-length(vec)), card))
  }
  
  deck(vec)
}
```

## Deck functions

Functions that we can apply to decks.

```{r}
# shuffle deck
shuffle <- \(d) {
  stopifnot(is.deck(d))
  d |> 
    sample(length(d), replace=F) |> 
    deck()
}

# draw top `size` cards from deck
draw <- \(d, size=7) {
  stopifnot(is.deck(d))
  deck(d[1:size])
}
```

## Counting functions

Functions to count colored sources.
This is where the juice of object oriented setup gets squeezed!

```{r}
size <- function(d) {
  stopifnot(is.deck(d))
  attr(d, "size")
}
count_spells <- function(d) {
  stopifnot(is.deck(d))
  attr(d, "size") - attr(d, "lands")
}
count_lands <- function(d) {
  stopifnot(is.deck(d))
  attr(d, "lands")
}

# count color sources in hand
count_colors <- function(d) {
  stopifnot(is.deck(d))
  
  COLORS |> 
    sapply(\(color) {
      d |> 
        sapply(\(card) {color %in% attr(card, "colors")}) |> 
        sum()
    })
}
```



# Example

```{r}
d <- build_deck(W=8, U=8, WUT=1)
d

d |> shuffle() |> draw() |> count_colors()
```


# TODO

* Implement "count untapped lands by turn x"
* Implement terramorphic expanse
* Implement MDFC



